Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-09-25 main.tex
Finished editing methods.

\subsection{Statistics}
%
We test our hypothesis that the balance assistance controller will reduce the
probability of falling when perturbed externally at the handlebar. We have a
single binary fall outcome variable \(f\) that is dependent on several possible
explanatory independent variables, one of which is the binary balance
assistance (on or off). See Table~\ref{tab:stat-model-variables} for the
statistical model variables.
%
\begin{table}
  \centering
  \caption{Independent and dependent variables used in the statistical model.}
  \label{tab:stat-model-variables}
  \begin{tabular}{llll}
    \toprule
    Variable & Causality & Units & Description \\
    \midrule
    \(L\)        & independent & \si{\newton\meter\second} & angular impulse of perturbation torque \\
    \(\delta_0\) & independent & \si{\degree} & steer angle at start of perturbation \\
    \(\phi_0\)   & independent & \si{\degree} & roll angle at start of perturbation \\
    \(f\)        & dependent   & boolean & outcome: did not fall 0 or did fall 1 \\
    \(c\)        & independent & boolean & balance assist: off 0 or on 1 \\
    \(j\)        & independent & integer & order number of perturbation \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Before fitting the model, we scale each independent variable \(x_{ij}^k\) such
that they have a mean of zero and a standard deviation of one by cluster-mean
centering, as recommended by \cite{Enders2007}, with clusters being an
individual participant. The clusters are chosen as all data from an individual
participant because we are only interested in the association between the state of
the balance assist system and the outcome of the perturbation. We expected
there to be a variation between participants in how well they are able to
resist the perturbation. However, this was not true. Cluster-mean centering
showed there to be no variation between participants. This fact allows us to
utilize a simple single-level logistic regression model, instead of a
multilevel model. This left us with angular impulse, perturbation order,
balance assist state, and roll \& steer angles at the time of perturbation as
independent variables. We also include interaction effects between the balance
assist state and the other four variables. We divide the analysis into two
separate model evaluations, one for the 6~\si{\kph}, \(g=10\) trials and one
for the 10~\si{\kph}, \(g=8\) trials and we evaluate our hypothesis for each
set of data. Statistics were computed with the lme4 version
1.1~\citep{Bates2015} package in R version 4.1.3~\citep{RCoreTeam2021}.

>>>>

\subsection{Statistics}
%
We test our hypothesis that the balance assist controller will reduce the
probability of falling when perturbed externally at the handlebar. We have a
single binary fall outcome variable \(f\) that is dependent on several possible
explanatory independent variables, one of which is the binary balance assist
state (on or off). See Table~\ref{tab:stat-model-variables} for the statistical
model variables.
%
\begin{table}
  \centering
  \caption{Independent and dependent variables used in the statistical
    regression model.}
  \label{tab:stat-model-variables}
  \begin{tabular}{llll}
    \toprule
    Variable & Causality & Units & Description \\
    \midrule
    \(L\)        & independent & \si{\newton\meter\second} & angular impulse of perturbation torque \\
    \(\delta_0\) & independent & \si{\degree} & steer angle at start of perturbation \\
    \(\phi_0\)   & independent & \si{\degree} & roll angle at start of perturbation \\
    \(f\)        & dependent   & boolean & outcome: did not fall 0 or did fall 1 \\
    \(c\)        & independent & boolean & balance assist: off 0 or on 1 \\
    \(j\)        & independent & integer & order number of perturbation \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Before fitting the model, we scale each independent variable \(x_{ij}^k\) such
that they have a mean of zero and a standard deviation of one by cluster-mean
centering, as recommended by \citet{Enders2007}, with clusters being an
individual participant. The clusters are chosen as all data from an individual
participant because we are only interested in the association between the state
of the balance assist system and the outcome of the perturbation. We expected
there to be a variation between participants in how well they are able to
resist the perturbation. However, this was not true. Cluster-mean centering
showed there to be no variation between participants. This fact allows us to
utilize a simple single-level logistic regression model, instead of a
multilevel model. This left us with angular impulse, perturbation order,
balance assist state, and roll \& steer angles at the time of perturbation as
independent variables. We also include interaction effects between the balance
assist state and the other four variables. We divide the analysis into two
separate model evaluations, one for the 6~\si{\kph}, \(g=10\) trials and one
for the 10~\si{\kph}, \(g=8\) trials and we evaluate our hypothesis for each
set of data. Statistics were computed with the lme4 version
1.1~\citep{Bates2015} package in R version 4.1.3~\citep{RCoreTeam2021}.
